---
title: "f1_analysis"
author: "Muntasir Adnan"
date: "2023-10-09"
output: pdf_document
---
```{r}
options(warn = -1)
```

```{r}
library(tidyverse)
library(sf)
# had to install os pacakage libproj-dev
# apt-get -y install libgdal-dev
# sudo apt-get install libudunits2-dev
library(mapview)
```

```{r}
circuits <- read.csv("data/circuits.csv")
constructors <- read.csv("data/constructors.csv")
```

```{r}
head(circuits)
```
```{r}
mapview(circuits, 
        xcol = "lng", 
        ycol = "lat", 
        crs = 4269, 
        label = paste("Circuit name - ", circuits$name,", Location - ", circuits$location, ", ", circuits$country, sep = ""), 
        # label = htmltools::HTML(paste(circuits$name, circuits$location, sep = '<br/>')),
        grid = FALSE)
```
Analyse the circuits
Starting with how many races those circuits had.
```{r}
races <- read.csv("data/races.csv")
```
change name to circuit_name in circuits
```{r}
names(circuits)[3] <- "circuit_name"
```
inner join
checking how many grand prix in each circuit and sorted in descending order
```{r}
races %>%
  inner_join(circuits, by = "circuitId") %>% 
  select(circuit_name, raceId, year, name) %>% 
  count(circuit_name) %>% 
  rename(count = n) %>%
  arrange(desc(count))
```
In total there were 74 grand prix events
```{r}
length(unique(races$year))
```
1 grand prix did not have monza, lets check which one
Check circuit id
```{r}
circuits %>% 
  filter(circuit_name %in% "Autodromo Nazionale di Monza") %>% 
  select(circuitId)
```
filtering all years that had monza
Checking which year monza was missing
```{r}
monza_gp <- races %>% 
  filter(circuitId == 14) %>% 
  select(year)
setdiff(unique(races$year), monza_gp$year)
```
2022 analysis
```{r}
f1_22_races <- races %>% 
  filter(year == 2022)
f1_22_races
```
2022 drivers' stanidings
race 1096 has all accumulated values
so last race of each season
```{r}
driver_standings <- read_csv("data/driver_standings.csv")
drivers <- read.csv("data/drivers.csv")

drivers <- drivers %>%
  unite(driver_name, forename, surname, sep = " ")
```
```{r}
f1_22_driver_standings <- driver_standings %>% 
  filter(raceId == 1096) %>% 
  inner_join(drivers, by = "driverId") %>% 
  select(number, driver_name, points, wins) %>% 
  arrange(desc(points))
f1_22_driver_standings
```
```{r}
library(ggplot2)
f1_22_driver_standings$driver_name <- reorder(f1_22_driver_standings$driver_name, f1_22_driver_standings$points)

ggplot(f1_22_driver_standings, 
       aes(x = f1_22_driver_standings$points, 
           y = f1_22_driver_standings$driver_name, fill = wins)) +
       geom_bar(stat = "identity") +
       labs(title = "2022 Drivers' Standings",
          x = "Points",
          y = "Driver Name",
          fill = "Wins") +
       theme(
          text = element_text(color = "Black", family = "Formula1 Display-Regular"),
          axis.text = element_text(color = "Black", size = 12, family = "Formula1 Display-Regular"),
          axis.title = element_text(color = "Black", size = 15, family = "Formula1 Display-Regular"), 
          title = element_text(size = 15, family = "Formula1 Display-Regular"),
          panel.grid = element_blank(),
          axis.ticks = element_blank(),
          panel.background = element_rect(fill = "white"),
          plot.background = element_rect(fill = "white"),
          plot.title = element_text(hjust = 0.5),
          axis.text.x = element_text(angle = 70,hjust = 1))+
       scale_fill_gradient(low = "tomato", high = "darkred")

```
pitstop analysis
```{r}
pit_stops <- read.csv("data/pit_stops.csv")
```
inner_join
```{r}
pit_with_year <- inner_join(races, pit_stops, by = "raceId") %>% 
  select(year, raceId, driverId, stop, lap, milliseconds)
```

```{r}
temp <- pit_with_year %>%
  group_by(year) %>%
  summarize(fastest_pit = min(milliseconds))
```
punish
```{r}
temp <- temp %>% 
  mutate(adj_time = ifelse(year<2014, fastest_pit*1, ifelse(year>=2014 & year<=2021, fastest_pit * 0.8, fastest_pit*0.6)))
# 2014-2021 20% reduce
# temp <- target %>%
#   filter(year >= 2014 & year <= 2021) %>%
#   mutate(fastest_lap = fastest_lap * 0.8)
# 
# # 2022 and 2023 40%
# temp2 <- target %>%
#   filter(year >= 2022 & year <= 2023 ) %>%
#   mutate(fastest_lap = fastest_lap * 0.6)
# 
# target <- bind_rows(
#   target %>% filter(year < 2014), # data before 2014
#   temp,
#   temp2
# )

  
```

```{r}
# ggplot(temp, aes(x = as.factor(year), y = fastest_pit, group = 1)) +
#   geom_line() +
#   labs(x = "Year", y = "Pit Time")
library(ggplot2)

# Assuming 'temp' is your data frame
ggplot(temp, aes(x = as.factor(year), y = fastest_pit, group = 1)) +
  geom_line(aes(color = "Fastest Pit Time")) +
  geom_line(data = temp, aes(x = as.factor(year), y = adj_time, color = "Adjusted Time")) +
  labs(x = "Year", y = "Pit-time") +
  scale_color_manual(values = c("Fastest Pit Time" = "blue", "Adjusted Time" = "red"))


```
Doing driver comparison


